Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-02-06 text/paper/revision_01/paper.tex
finalized edits sections 1-3 K2 revision 01

\begin{abstract}
  The scalar item \emph{some} is widely assumed to receive a pragmatic
  meaning enrichment to \emph{some but not all} if it occurs in matrix
  position. The question under which circumstances this enrichment can
  occur in certain embedded positions plays an important role in
  deciding how to delineate semantics and pragmatics. % Some claim
  % that it requires marked intonation for the relevant pragmatic
  % enrichment to occur in embedded position; others allow for it more
  % freely.
  We present new experimental data that bear on this theoretical
  issue. In distinction to previous experimental approaches, we
  presented sentence material auditorily in order to explicitly
  control prosodic markedness of the scalar item. Moreover, our
  experiment sheds light on the relative preferences or salience of
  candidate readings. The presented data turn out to be problematic
  for both a traditional Gricean view of scalar implicature
  calculation, as well as the more recent grammatical approach.
\end{abstract}

@@

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings compared to
each other (if they are detectable by empirical means at all)?
Addressing this question empirically is relevant because it lies at
the heart of a current debate about the nature of quantity inferences.
The controversy is usually seen as one between two main competing
theoretical positions, which we call \mymark{traditionalism} and
\mymark{grammaticalism} and take to be general ideas behind large
classes of approach to quantity implicatures \citep[c.f.][for related
discussion with a focus on deriving processing
theories]{ChemlaSingh2014:Remarks-on-the-}. Unfortunately, it is
rather difficult to empirically test either ``core theory'' (
terminology from \citeauthor{ChemlaSingh2014:Remarks-on-the-})
directly because additional ``auxiliary assumptions'' are needed in
both cases to derive falsifiable \emph{ex ante} predictions about
to-be-expected observations in a concrete behavioral experiment, even
irrespective of processing accounts. We therefore try to carefully set
up plausible instantiations of either core theory with particular
auxiliary assumptions that give testable predictions at the outset in
Section~\ref{sec:theories-predictions}, then evaluate the obtained
predictions against our data, and finally broaden our perspective to
reflect in Section~\ref{sec:critical-reflection} on the impact of our
data on the more general tenability of the core theories.

But even with crisp theoretical predictions at hand, matters are still
very complex, as reflected by lively discussions about which
conclusions are permitted by previous experimental research
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. The
results of these studies were discussed controversially because of
several specific reasons. Firstly, and most importantly, previous
studies presented target sentences visually.  However, traditionalists
often acknowledge the availability of local readings for prosodically
marked utterances, albeit not as the result of a standard scalar
implicature but of a different pragmatic process triggered by prosodic
markedness
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. Under
this view, evidence for the availability of local readings
\citep[see][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}
may be an effect of “silent prosody” \citep[see
e.g.][]{Bader98,Fodor98}. We will therefore address this extant
position as the \mymark{prosodic markedness hypothesis} and argue
that, if prosody is hypothesized to play a central role, it is vital
to present target sentences auditorily or to control for effects of
“silent prosody” that subjects may apply when reading a
sentence. Secondly, as \citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences,
might well be confounded with, in particular, “typicality effects”
(see Section~\ref{sec:previous-studies} for discussion).  Finally,
previous studies have only accumulated limited evidence pertaining to
the relative salience of candidate readings of sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE}) (see
Section~\ref{sec:previous-studies} for discussion).

% More generally, two prerequisites have to be met in order to make use
% of experimental data to decide between theoretical positions regarding
% the computation of quantity inferences. Firstly, both traditionalism
% and grammaticalism have to be augmented with additional assumptions to
% derive testable predictions about the relative salience of candidate
% readings. For example, hypothezised effects of prosodic marking have
% to be taken into account before we can test traditionalism empirically
% using sentences (\ref{bsp:AE}) and (\ref{bsp:GE}).  We will address
% this extant position as the \mymark{prosodic markedness hypothesis}
% and argue that, if prosody is hypothesized to play a central role, it
% is vital to present target sentences auditorily or to control for
% effects of “silent prosody” that subjects may apply when reading a
% sentence. Secondly, the relative salience of candidate readings has to
% be operationalized properly, i.e., we have to devise methods to
% measure relative salience. For example, typicality effects and logical
% dependencies between readings have to be controlled for when
% evaluating speaker judgments in terms of the salience of candidate
% readings.  Both of these prerequisites are related to formulating what
% \citet{ChemlaSingh2014:Remarks-on-the-} call ``auxiliary assumptions''
% that are necessary to enable theoretical accounts to make predictions
% about processing. Our emphasis here is not on processing accounts, but
% on predictions of behavioral data --- what is sometimes, perhaps
% misleadingly, called a computational-level perspective. But even so,
% traditionalism and grammaticalism have to summon auxiliary assumptions
% which connect the core theories to the dependent variables of a
% behavioral experiment. Without proper auxiliary assumptions it is hard
% (or impossible) to test traditionalism or grammaticalism empirically.

@@

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out plausible
\emph{ex ante} predictions of the core theories traditionalism and
grammaticalism (for some practically relevant auxiliary assumptions)
about availability and relative salience of readings for our target
sentences. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} eventually contributes our
experimental design, analyses and results. Finally,
Section~\ref{sec:critical-reflection} discusses our method and
interpretation of the results in a broader theoretical context.

@@

There are two main theoretical positions which appear to make
different predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \mymark{traditionalism} and
\mymark{grammaticalism} and treat each in turn. Both approaches are
addressed here as ``core theories'': two main ideas behind several
approaches to quantity implicatures. To determine concrete predictions
from these, additional ``auxiliary assumptions'' are necessary. Here,
we spell a set of such auxiliary assumptions for each core theory. Our
motivation for the particular choices we make is partly guided by
practical concerns, so as to have a concrete set of predictions for
each core theory to formulate expectations and compare our data to. We
therefore introduce some prominent or otherwise relevant variants of
either core theory, the predictions of which are summarized in
Table\ref{tab:predictions} on page~\pageref{tab:predictions}. How
other potential instantiations of either core theory fare in light of
our data will be discussed in Section~\ref{sec:critical-reflection}.

Although strictly speaking not a part of a Gricean theory of scalar
implicature calculation, traditionalists frequently concede that local
readings are in principle available, albeit not as a standard scalar
implicature, but as a different pragmatic process triggered by
prosodic markedness. We address this \emph{prosodic markedness
  hypothesis} separately in Section~\ref{sec:inton-mark-hypoth},
alongside a brief revision of the relevant experimental data.

@@

The traditionalist view gives rise to what
\citet{Geurts2010:Quantity-Implic} calls the \mymark{Gricean recipe}
for deriving a scalar inference (\ref{bsp:Plain-SI-Implicature}) from
an utterance of (\ref{bsp:Plain-SI-Target}): if the issue whether Hans
solved only some or all of the problems is relevant, then a
cooperative and knowledgable speaker would utter
(\ref{bsp:Plain-SI-Alternative}) if in a position to do so; hence, one
of the most natural explanations of why such a speaker has not uttered
(\ref{bsp:Plain-SI-Alternative}), but only (\ref{bsp:Plain-SI-Target})
is that she is uncertain of whether (\ref{bsp:Plain-SI-Alternative})
is true; but on the assumption that she is knowledgeable (competent,
opinionated, informed \dots) it follows that
(\ref{bsp:Plain-SI-Implicature}) should in fact be true.\fn{We are
  glossing here somewhat swiftly over the more nuanced details of the
  derivation of implicatures targeting the speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit}. More
  on this below.}

@@

Traditionalism's predictions about availability and relative salience
of readings of \as- and \es-sentences depend on further auxiliary
assumptions. To have handy labels, we distinguish between strong and
weak variants, and between restricted and unrestricted
variants.

Strong and weak traditionalism differ with respect to whether the
assumptions needed to derive non-epistemic implicatures are prominent
or not in the given context of utterance (or some experiment or
other). These assumptions include relevance of the information in
question, mutual awareness that the stronger alternative has been a
speaker option, the speaker's competence about the issue,
etc. \emph{Strong traditionalism} assumes that the additional
assumptions needed to derive a non-epistemic implicatures are
sufficiently salient to prefer the non-epistemic implicature over the
literal reading; \emph{weak traditionalism} assumes the reverse
preference order. In other words, in the context of a given
experiment, it is compatible with a traditionalist position to expect
literal or implicature readings to be more prominent. Or, put yet
differently, a traditionalist could be uncertain as to whether a given
experiment is more likely to trigger more implicature readings or
literal readings. Without more information about the experiment or
context of interpretation, traditionalism is compatible with either
salience ordering.

Orthogonal to this, we distinguish restricted and unrestricted
variants of traditionalism. For our purposes, restricted
traditionalism considers only alternative expressions that are
stronger than the to-be-interpreted utterance. Unrestricted
traditionalism is more relaxed and also allows alternatives that are
derived from replacing \emph{some} with \emph{all} and vice versa, if
necessary multiple times, even if the resulting alternative is not
logically stronger. Restricted traditionalism considers only
\as-sentences to have a relevant alternative, namely
(\ref{bsp:AE-Alternative}). It then predicts the availability of a
global reading for \as-sentences on top of the literal. Restricted
traditionalism predicts only literal readings of \es-sentences,
because the alternative (\ref{bsp:GE-Alternative}) does not entail the
literal reading of (\ref{bsp:GE}). Every implicature reading available
to restricted traditionalism is also available to unrestricted
traditionalism, but the latter's predictions are even richer. For
\as-sentences, the potential alternative
(\ref{bsp:AE-Alternatives-Extended}) also makes local readings
available \citep[e.g.][]{Sauerland2004:Scalar-Implicat}: conjoining a
literal reading of (\ref{bsp:AE}) with the negation of
(\ref{bsp:AE-Alternatives-Extended}) gives a local reading.

@@

\noindent For \es-sentences, we derive a global reading via the
negation of alternative (\ref{bsp:GE-Alternative}).

In sum, a traditionalist core theory is compatible with a range of
observations about the availability and relative salience of
implicature readings. These are summarized in
Table\ref{tab:predictions}. Most importantly for the subsequent
discussion, traditionalism does not predict local readings for
\es-sentences and can favor implicature readings over literal ones or
vice versa.

@@

The central idea behind grammaticalist approaches is that quantity
implicatures are derived by application of exhaustivity operators at
varying scope-sites during compositional computation of a sentence's
truth-value, if necessary multiple times
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Fox2007:Free-Choice-and,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. For
our purposes, it is enough to assume that $\exh(\cdot)$ is a poly-typed
function that enriches an expression $X$, which need not be a full
proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:\fn{More sophisticated formulations of exhaustive interpretation
  have been proposed
  \citep[e.g.][]{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}
%
A grammaticalist core theory predicts that all three readings of \as-
and \es-sentences are in principle available: literal readings arise
if no exhaustification operator is applied; global readings arise if
the exhaustification operator takes sentence-wide scope as in
(\ref{Grammar-Global}); and local readings arise if the
exhaustification operator takes scope under the respective quantifiers
as in (\ref{Grammar-Local}).

@@

It is a topic of active current debate as to how a grammaticalist
approach should be supplemented with a principled \emph{disambiguation
  mechanism} so as to rule out unattested readings obtainable by
embedded exhaustification and to make concrete predictions about
preferences over readings made available. % Grammaticalists therefore
% acknowledge that \emph{some} disambiguation principle must be
% specified, for otherwise grammaticalism would hardly make any
% commitment and would overpredict readings in cases where they are
% unattested, i.e., when scalar items appear in downward-entailing
% contexts \citep[c.f.][for explicit
% discussion]{FoxSpector:Economy-and-Emb}.
Three kinds of approaches have been suggested in the literature
\citep[c.f.][]{ChemlaSingh2014:Remarks-on-the-}. We briefly touch upon
each, but focus on the one that is least controversially applicable to
our experimental case, and that has theoretical significance beyond
the debate about embedded implicatures.

To start with,
\citet{Magri2009:A-Theory-of-Ind,Magri2011:Another-Argumen} opts for
obligatory presence of $\exh(\cdot)$ with variability in the set of relevant
alternatives.  Under this approach, alternatives may be pruned
entirely, so that all occurrences of $\exh(\cdot)$ become vacuous resulting
in a literal reading. Whether alternatives apply or not is a matter of
their relevance. It is not clear what his approach predicts \emph{ex
  ante} about the relative salience of the candidate readings in our
experiment, but we will show in Section~\ref{sec:critical-reflection}
that this approach can be brought to accommodate the data well after
the fact.

@@

Luckily, the last kind of disambiguation approach does yield a clear
and non-trivial restriction of reading preferences off-the-shelf. It
is to select readings by comparison of logical strengths, as suggested
by the \mymark{strongest meaning hypothesis} of
\citet{DalrympleKanazawa1998:Reciprocal-Expr}. Supplementing
grammaticalism with a strength-based selection criterion has been
suggested, in one form or another, in many contributions to this
paradigm
\citep[e.g.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. At
the same time, the strongest meaning hypothesis has been frequently
called upon for disambiguation purposes in other domains as well,
including reciprocals \citep{DalrympleKanazawa1998:Reciprocal-Expr},
plural predication \citep{Winter2001:Plural-Predicat} or vagueness
\citep{CobrerosEgre2012:Tolerant-Classi}. For these reasons, we focus
from now on this latter type of selection principle. In the discussion
section (section~\ref{sec:discussion-1}) alternative selection
principles will be reconsidered.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation that make non-trivial predications
about preferred readings of \as- and \es-sentences (unlike, say, the
more detailed variants of \cite{FoxSpector:Economy-and-Emb}): fix a
sentence with propositional content $S$ and candidate readings $C(S)$,
where $C(S)$ contains $S$ and all readings derivable from inserting
$\exh(\cdot)$ at suitable scope sites;\fn{Strictly speaking, we should
  compare \emph{parses} of sentences with $\exh(\cdot)$ at various
  scope sites with respect to the readings that these parses give rise
  to \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may
  ignore this detail here for ease of exposition.} then define for
$X,Y \in C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\mymark{weak grammaticalism} and \mymark{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

Taking stock, traditionalism and grammaticalism have been introduced
as fairly general core theories that need backup from particular
auxiliary assumptions to make concrete predictions about availability
and readings preferences for our target sentences. Most notably, under
the assumptions made here, traditionalism does not predict local
readings for \es-sentences to be available while grammaticalism does.
Grammaticalism, when supplemented with the the strongest meaning
hypothesis as disambiguation principle, predicts that all candidate
readings are available and that literal readings are dispreferred in
all cases. All predictions are summarized in
Table~\ref{tab:predictions}.

>>>>

\begin{abstract}
  The scalar item \emph{some} is widely assumed to receive a pragmatic
  meaning enrichment to \emph{some but not all} if it occurs in matrix
  position. The question under which circumstances this enrichment can
  occur in certain embedded positions plays an important role in
  deciding how to delineate semantics and pragmatics. % Some claim
  % that it requires marked intonation for the relevant pragmatic
  % enrichment to occur in embedded position; others allow for it more
  % freely.
  We present new experimental data that bear on this theoretical
  issue. In distinction to previous experimental approaches, we
  presented sentence material auditorily in order to explicitly
  control prosodic markedness of the scalar item. Moreover, our
  experiment sheds light on the relative preferences or salience of
  candidate readings. The presented data turn out to be unexpected
  under a traditional Gricean view, but also suggest refinement of the
  more recent grammatical approach.
\end{abstract}

@@

Given this plurality of hypothesized readings, the question arises:
how salient are these theoretically conceivable readings compared to
each other (if they are detectable by empirical means at all)?
Addressing this question empirically is relevant because it lies at
the heart of a current debate about the nature of quantity inferences.
The controversy is usually seen as one between two main competing
theoretical positions, which we call \emph{traditionalism} and
\emph{grammaticalism} and take to be general ideas behind large
classes of approach to quantity implicatures \citep[c.f.][for related
discussion with a focus on deriving processing
theories]{ChemlaSingh2014:Remarks-on-the-}. Unfortunately, it is
rather difficult to empirically test either ``core theory''
(terminology from \citeauthor{ChemlaSingh2014:Remarks-on-the-})
directly because additional ``auxiliary assumptions'' are needed to
derive falsifiable \emph{ex ante} predictions about to-be-expected
observations in a concrete behavioral experiment, even irrespective of
processing accounts. We therefore try to carefully set up plausible
instantiations of either core theory with particular auxiliary
assumptions that give testable predictions at the outset in
Section~\ref{sec:theories-predictions}, then evaluate the obtained
predictions against our data, and finally broaden our perspective to
reflect in Section~\ref{sec:critical-reflection} on the impact of our
data on the more general tenability of the core theories.

But even with crisp theoretical predictions at hand, matters are still
very complex, as reflected by lively discussions about which
conclusions are permitted by previous experimental research
\citep[e.g.][]{GeurtsPouscoulous2009:Embedded-Implic,CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev,BenzGotzner2014:Embedded-implic}. The
results of these studies were discussed controversially because of
several specific reasons. Firstly, and most importantly, previous
studies presented target sentences visually.  However, traditionalists
often acknowledge the availability of local readings for prosodically
marked utterances
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar,GeurtsTielvan-Tiel2013:Embedded-Scalar}. Under
this view, evidence for the availability of local readings
\citep[see][]{CliftonDube2010:Embedded-Implic,ChemlaSpector2010:Experimental-Ev}
may be an effect of “silent prosody” \citep[see
e.g.][]{Bader98,Fodor98}. We therefore address this position as the
\emph{prosodic markedness hypothesis} explicitly. Secondly, as
\citet{Tielvan-Tiel2012:Embedded-Scalar} and
\citet{GeurtsTielvan-Tiel2013:Embedded-Scalar} argue, the alleged
evidence for the availability of local readings of \as-sentences,
might well be confounded with, in particular, “typicality effects”
(see Section~\ref{sec:previous-studies} for discussion).  Finally,
previous studies have only accumulated limited evidence pertaining to
the relative salience of candidate readings of sentences like
(\ref{bsp:AE}) and (\ref{bsp:GE}) (see
Section~\ref{sec:previous-studies} for discussion).

% More generally, two prerequisites have to be met in order to make
% use of experimental data to decide between theoretical positions
% regarding the computation of quantity inferences. Firstly, both
% traditionalism and grammaticalism have to be augmented with
% additional assumptions to derive testable predictions about the
% relative salience of candidate readings. For example, hypothezised
% effects of prosodic marking have to be taken into account before we
% can test traditionalism empirically using sentences (\ref{bsp:AE})
% and (\ref{bsp:GE}).  We will address this position as the
% \mymark{prosodic markedness hypothesis} and argue that, if prosody
% is hypothesized to play a central role, it is vital to present
% target sentences auditorily or to control for effects of “silent
% prosody” that subjects may apply when reading a sentence. Secondly,
% the relative salience of candidate readings has to be
% operationalized properly, i.e., we have to devise methods to measure
% relative salience. For example, typicality effects and logical
% dependencies between readings have to be controlled for when
% evaluating speaker judgments in terms of the salience of candidate
% readings.  Both of these prerequisites are related to formulating
% what \citet{ChemlaSingh2014:Remarks-on-the-} call ``auxiliary
% assumptions'' that are necessary to enable theoretical accounts to
% make predictions about processing. Our emphasis here is not on
% processing accounts, but on predictions of behavioral data --- what
% is sometimes, perhaps misleadingly, called a computational-level
% perspective. But even so, traditionalism and grammaticalism have to
% summon auxiliary assumptions which connect the core theories to the
% dependent variables of a behavioral experiment. Without proper
% auxiliary assumptions it is hard (or impossible) to test
% traditionalism or grammaticalism empirically.

@@

The paper is structured as follows. Section~\ref{sec:get-know-your}
elaborates on the three kinds of relevant readings for our target
sentences. Section~\ref{sec:theories-predictions} works out plausible
\emph{ex ante} predictions of the competing core
theories. Section~\ref{sec:previous-studies} reviews related
experimental studies to the extent necessary to motivate our own
approach. Section~\ref{sec:exp} describes our experimental design,
analyses and results. Finally, Section~\ref{sec:critical-reflection}
discusses our method and interpretation of the results in a broader
theoretical context.

@@

There are two main theoretical positions which appear to make
different predictions about the readings of \as- and \es-sentences
\citep[see][for
overview]{Horn2006:The-Border-Wars,Geurts2010:Quantity-Implic,Sauerland2012:The-Computation,ChemlaSingh2014:Remarks-on-the-}. We
will refer to these as \emph{traditionalism} and \emph{grammaticalism}
and treat each in turn. Both approaches are addressed here as ``core
theories'': two main ideas behind several approaches to quantity
implicatures. To determine concrete predictions from these, additional
``auxiliary assumptions'' are necessary. Here, we spell a set of such
auxiliary assumptions for each core theory. Our motivation for the
particular choices we make is partly guided by practical concerns, so
as to have a concrete set of predictions for each core theory to
formulate expectations and compare our data to. We therefore introduce
some prominent or otherwise relevant variants of either core theory,
the predictions of which are summarized in Table~\ref{tab:predictions}
on page~\pageref{tab:predictions}. How other potential instantiations
of either core theory fare in light of our data will be discussed in
Section~\ref{sec:critical-reflection}.

Traditionalists frequently concede that local readings are licensed if
prosodically marked. We address this \emph{prosodic markedness
  hypothesis} separately in Section~\ref{sec:inton-mark-hypoth},
alongside a brief revision of the relevant experimental data.

@@

The traditionalist view gives rise to what
\citet{Geurts2010:Quantity-Implic} calls the \emph{Gricean recipe} for
deriving (\ref{bsp:Plain-SI-Implicature}) from
(\ref{bsp:Plain-SI-Target}): if the issue whether Hans solved only
some or all of the problems is relevant, then a cooperative and
knowledgable speaker would utter (\ref{bsp:Plain-SI-Alternative}) if
in a position to do so; hence, one of the most natural explanations of
why such a speaker has not uttered (\ref{bsp:Plain-SI-Alternative}),
but only (\ref{bsp:Plain-SI-Target}) is that she is uncertain of
whether (\ref{bsp:Plain-SI-Alternative}) is true; but on the
assumption that she is knowledgeable (competent, opinionated, informed
\dots) it follows that (\ref{bsp:Plain-SI-Implicature}) should in fact
be true.\fn{We are glossing here somewhat swiftly over the more
  nuanced details of the derivation of implicatures targeting the
  speaker's epistemic state
  \citep[e.g.][]{Gazdar1979:Pragmatics:-Imp,Soames1982:How-Presupposit}. More
  on this below.}

@@

Traditionalism's predictions about availability and relative salience
of readings of \as- and \es-sentences depend on further auxiliary
assumptions. To have handy labels, we distinguish between strong and
weak, and between restricted and unrestricted variants.

Strong and weak traditionalism differ with respect to whether the
assumptions needed to derive non-epistemic implicatures are prominent
in the given context of utterance (or some experiment or other). These
assumptions include relevance of the information in question, mutual
awareness that the stronger alternative has been a speaker option, the
speaker's competence about the issue, etc. \emph{Strong
  traditionalism} assumes that the additional assumptions needed to
derive non-epistemic implicatures are sufficiently salient to prefer
an implicature reading over a literal reading; \emph{weak
  traditionalism} assumes the reverse preference order. In other
words, in the context of a given experiment, it is compatible with a
traditionalist position to expect literal or implicature readings to
be more prominent. Or, put yet differently, a traditionalist could be
uncertain as to whether a given experiment is more likely to trigger
more implicature readings or literal readings. Without more
information about the experiment or context of interpretation,
traditionalism is compatible with either salience ordering.

Orthogonal to this, we distinguish restricted and unrestricted
variants of traditionalism. For our purposes, restricted
traditionalism considers only alternative expressions that are
stronger than the to-be-interpreted utterance. Unrestricted
traditionalism also allows alternatives that are derived from
replacing \emph{some} with \emph{all} and vice versa, if necessary
multiple times, even if the resulting alternative is not logically
stronger. Restricted traditionalism considers only \as-sentences to
have a relevant alternative, namely (\ref{bsp:AE-Alternative}). It
then predicts the availability of a global reading for \as-sentences
on top of the literal. Restricted traditionalism predicts only literal
readings of \es-sentences, because the alternative
(\ref{bsp:GE-Alternative}) does not entail the literal reading of
(\ref{bsp:GE}). Every implicature reading available to restricted
traditionalism is also available to unrestricted traditionalism, but
the latter's predictions are even richer. For \as-sentences, the
potential alternative (\ref{bsp:AE-Alternatives-Extended}) also makes
local readings available
\citep[e.g.][]{Sauerland2004:Scalar-Implicat}: conjoining a literal
reading of (\ref{bsp:AE}) with the negation of
(\ref{bsp:AE-Alternatives-Extended}) gives a local reading.

@@

\noindent For \es-sentences, we derive a global reading via the
negation of alternative (\ref{bsp:GE-Alternative}), but not local
readings.

In sum, a traditionalist core theory is compatible with a range of
observations about the availability and relative salience of
implicature readings. These are summarized in
Table~\ref{tab:predictions}. Most importantly for the subsequent
discussion, traditionalism does not predict local readings for
\es-sentences and can favor implicature readings over literal ones or
vice versa.

@@

The central idea behind grammaticalist approaches is that quantity
implicatures are derived by application of exhaustivity operators at
varying scope-sites during compositional computation of a sentence's
truth-value
\citep[see][]{Chierchia2006:Broaden-Your-Vi,Magri2011:Another-Argumen,Sauerland2012:The-Computation,ChierchiaFox2008:The-Grammatical}. For
our purposes, it is enough to assume that $\exh(\cdot)$ is a
poly-typed function that enriches an expression $X$, which need not be
a full proposition, based on a set $\alt(X)$ of (suitable, relevant)
alternatives to $X$ that yields an enriched meaning of the
form:% \fn{More sophisticated formulations of exhaustive interpretation
  % have been proposed
  % \citep[e.g.][]{GroenendijkStokhofThesis1984,Stechowvon-StechowZimmermann1984:Term-Answers-an,Schulz2005:A-Pragmatic-Sol,vanRooijSchulz:ExhaustiveInterpretation,Spector2006:Scalar-Implicat,Fox2007:Free-Choice-and}.}
\begin{exe}
  \ex \label{bsp:Exh-Def} $\exh(X,\alt(X)) = X \bigwedge_{A \in
      \alt(X)} \neg A$.
\end{exe}
%
A grammaticalist core theory predicts that all three readings of \as-
and \es-sentences are in principle available: literal readings arise
if no exhaustification operator is applied; global readings arise if
the exhaustification operator takes sentence-wide scope as in
(\ref{Grammar-Global}); and local readings arise if the
exhaustification operator takes scope under the respective quantifiers
as in (\ref{Grammar-Local}).

@@

It is a topic of active current debate as to how a grammaticalist
approach should be supplemented with a general and well-motivated
\emph{disambiguation criterion} so as to rule out unattested readings
obtainable by embedded exhaustification and to make concrete
predictions about preferences over readings made
available. % Grammaticalists therefore
% acknowledge that \emph{some} disambiguation principle must be
% specified, for otherwise grammaticalism would hardly make any
% commitment and would overpredict readings in cases where they are
% unattested, i.e., when scalar items appear in downward-entailing
% contexts \citep[c.f.][for explicit
% discussion]{FoxSpector:Economy-and-Emb}.
Three kinds of approaches have been suggested in the literature
\citep[c.f.][]{ChemlaSingh2014:Remarks-on-the-}. We briefly touch upon
each, but focus on the one that is least controversially applicable to
our experimental case, and that has theoretical significance beyond
the debate about embedded implicatures.

To start with,
\citet{Magri2009:A-Theory-of-Ind,Magri2011:Another-Argumen} opts for
obligatory presence of $\exh(\cdot)$ with variability in the set of
alternatives to feed into each $\exh(\cdot)$. Under this approach,
alternatives may be pruned entirely, so that all occurrences of
$\exh(\cdot)$ become vacuous, resulting in a literal reading. Whether
alternatives apply or not is a matter of their relevance. It is not
clear what his approach predicts \emph{ex ante} about the relative
salience of candidate readings in a neutral experiment that does not
explicitly provide information about which alternatives are relevant
(for whatever purpose). However, we will show in
Section~\ref{sec:critical-reflection} that Magri's approach can be
brought to accommodate the data very well after the fact.

@@

Luckily, the last kind of disambiguation approach does yield a clear
and non-trivial restriction of reading preferences off-the-shelf. It
selects readings by comparison of logical strengths, as suggested by
the \emph{strongest meaning hypothesis} of
\citet{DalrympleKanazawa1998:Reciprocal-Expr}. Supplementing
grammaticalism with a strength-based selection criterion has been
suggested, in one form or another, in many contributions to this
paradigm
\citep[e.g.][]{FoxSpector:Economy-and-Emb,ChierchiaFox2008:The-Grammatical,Chierchia2012:FC-Nominals-and,Spector2014:Global-Positive}. At
the same time, the strongest meaning hypothesis has been frequently
called upon for disambiguation purposes in other domains as well,
including reciprocals \citep{DalrympleKanazawa1998:Reciprocal-Expr},
plural predication \citep{Winter2001:Plural-Predicat} or vagueness
\citep{CobrerosEgre2012:Tolerant-Classi}. For these reasons, we focus
from now on this latter type of selection
principle. Section~\ref{sec:critical-reflection} reconsiders
alternative selection principles.

\citet{ChierchiaFox2008:The-Grammatical} discuss two concrete variants
of strength-based disambiguation that make non-trivial predications
about preferred readings of \as- and \es-sentences (unlike, say, the
more detailed variants of \cite{FoxSpector:Economy-and-Emb}): fix a
sentence with propositional content $S$ and candidate readings $C(S)$,
where $C(S)$ contains $S$ and all readings derivable from inserting
$\exh(\cdot)$ at suitable scope sites;\fn{Strictly speaking, we should
  compare \emph{parses} of sentences with $\exh(\cdot)$ at various
  scope sites with respect to the readings that these parses give rise
  to \citep[see][]{ChierchiaFox2008:The-Grammatical}, but we may
  ignore this detail here for ease of exposition.} then define for
$X,Y \in C(S)$ either
\begin{exe}
\ex \label{bsp:SMH-Total} 
  $X$ is preferred to $Y$ iff $X \subset Y$ 
  \attrib{\citep[(104)]{ChierchiaFox2008:The-Grammatical}}
\end{exe}
or
\begin{exe}
\ex \label{bsp:SMH-Partial}  $X$ is preferred to $Y$ iff $X \subset Y$
  and $X$ and $Y$ are readings obtained from applications of
  $\exh(\cdot)$ at exactly the same scope sites, except for
  one. \attrib{\citep[(105)]{ChierchiaFox2008:The-Grammatical}} 
\end{exe}
Since the former is entailed by the latter, we will speak of
\emph{weak grammaticalism} and \emph{strong grammaticalism},
depending on whether (\ref{bsp:SMH-Total}) or (\ref{bsp:SMH-Partial})
is used.

@@

Taking stock, traditionalism and grammaticalism have been introduced
as fairly general core theories that need backup from particular
auxiliary assumptions to make concrete predictions about availability
and reading preferences for our target sentences. Most notably, under
the assumptions made here, traditionalism does not predict local
readings for \es-sentences to be available while grammaticalism does.
Grammaticalism, in the form proposed by
\citet{ChierchiaFox2008:The-Grammatical}, predicts that all candidate
readings are available and that literal readings are dispreferred in
all cases. All predictions are summarized in
Table~\ref{tab:predictions}.
